A Hitchhiker's Guide to Reproducible Research in R

### 2021-06-15 manuscript.Rmd
address andreas' comments

Increasingly, scientists openly share their data, materials, and analysis code.
Sharing these digital artefacts increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results.
However, making a research project reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.

This paper has two aims.
First, we want to introduce the reader to a workflow [see @Peikert2019 for details] that ensures reproducibility---defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code. 
The research community has long accepted reproducibility as integral to "good scientific practice" because empirical research hinges on its objectivity meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22][^popper]
Often, however, projects can not be reproduced by other research teams, despite the best efforts by the creators.
We present four common threats to reproducibility and explain how we can address these with a combination of four solutions from the software engineering community.
Specifically, we demonstrate how the R package `repro` supports researchers in implementing this solution.
The second objective is to present an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the bulk of the manuscript, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.

@@

Even though transparency has increased across scholarly disciplines (that is, data and methods are increasingly openly shared), most of these open repositories do not provide sufficient information to reproduce relevant computational and statistical results.
These failed attempts to reproduce highlight the need for widely accepted reproducibility standards.
To increase the proportion of reproducible projects, such reproducibility standard has to balance the trade-off between being rigorous and being easy to use.
A rigorous standard increases the likelihood of any individual project to be actually reproducible.
An easy to use standard, on the other hand, is more likely to be employed across many research projects.
The remainder of this paper is structured along these lines.
The first part introduces theoretical concepts and software solutions that together form a rigorous reproducibility standard.
The second part is a tutorial that shows how these software solutions can be easily employed by researchers with the help of the R package `repro`.

@@

> "Is there a mean difference in the personality trait 'Machiavellism' between genders?"

@@

To automatically resolve dependencies between project files, we rely on **dependency tracking**.
In essence, developers provide a collection of computational recipes that describe how ingredients are processed to create intermediate products and, in the end, a final product.
Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, analysis results).

@@

A significant hurdle for adopting the before-mentioned tools for researchers is that they are typically used through a command-line interface (CLI).
The CLI is vastly more user friendly than its predecessor, the punch card, but many users (including the authors) prefer a graphical user interface (GUI) over text-based interactions for most tasks.
A GUI has the apparent disadvantages for reproducibility that it requires a user who clicks on the correct buttons in the correct order.
Depending on the number of steps, this is not only much effort but also error-prone.
This tutorial seeks a compromise.
We use simpler alternatives, including a GUI where we can, to create a reproducible workflow, but the reproduction of such workflow does not rely on manual steps.
At several points, we rely on or have self-developed a CLI that wraps a more complicated (and powerful) CLI.
These wrappers use standard CLI tools but provide feedback for the user about what the computer does and what the user should do in layman's terms and are accessible from within R.
We hope this makes reproducibility tools more accessible by enabling untrained users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
These alternative tools are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.
Please install the package [`gert`](https://github.com/r-lib/gert)[@R-gert] which wraps Git commands and [`repro`](https://github.com/aaronpeikert)[@R-repro] so they can act as assistance systems while you follow the tutorial:

>>>>

Increasingly, scientists openly share their data, materials, and analysis code.
Sharing these digital artefacts increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results, but only if these artifacts are reproducible.
However, making a research project reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.

This paper has two aims.
First, we want to introduce the reader to a workflow [see @Peikert2019 for details] that ensures reproducibility---defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code. 
The research community has long accepted reproducibility as integral to "good scientific practice" because empirical research hinges on its objectivity meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22][^popper]
Often, however, projects can not be reproduced by other research teams, despite the best efforts by the creators.
We present four common threats to reproducibility and explain how we can address these with a combination of four solutions from the software engineering community.
Specifically, we demonstrate how the R package `repro` supports researchers in implementing this solution.
The second objective is to present an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.

@@

These failed attempts to reproduce highlight the need for widely accepted reproducibility standards.
These open repositories do not routinely provide sufficient to reproduce relevant computational and statistical results.
We argue that the research community could benefit more from the increased transparency across scholarly disciplines (that is, data and methods are increasingly openly shared), if more projects could be reproduced.
To increase the proportion of reproducible projects, such reproducibility standard has to balance the trade-off between being rigorous and being easy to use.
A rigorous standard increases the likelihood of any individual project to be actually reproducible.
An easy to use standard, on the other hand, is more likely to be employed across many research projects.
The remainder of this paper is structured along these lines.
The first part introduces theoretical concepts and software solutions that together form a rigorous reproducibility standard.
The second part is a tutorial that shows how these software solutions can be easily employed by researchers with the help of the R package `repro`.

@@

> "Is there a mean difference in the personality trait 'Machiavellism' between self identified females and males?"

@@

To automatically resolve dependencies between project files, we rely on **dependency tracking**.
In essence, researchers provide a collection of computational recipes that describe how ingredients are processed to create intermediate products and, in the end, a final product.
Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, analysis results).

@@

A significant hurdle for adopting the before-mentioned tools for researchers is that they are typically used through a command-line interface (CLI).
The CLI predates and coexists with graphical user interfaces (GUIs) which many users (including the authors) prefer over the text-based interactions of a CLI for most tasks.
A GUI has the apparent disadvantages for reproducibility that it requires a user who clicks on the correct buttons in the correct order.
Depending on the number of steps, this is not only much effort but also error-prone.
This tutorial seeks a compromise.
We use simpler alternatives, including a GUI where we can, to create a reproducible workflow, but the reproduction of such workflow does not rely on manual steps.
At several points, we rely on or have self-developed a CLI that wraps a more complicated (and powerful) CLI.
These wrappers use standard CLI tools but provide feedback for the user about what the computer does and what the user should do in layman's terms and are accessible from within R.
We hope this makes reproducibility tools more accessible by enabling untrained users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
These alternative tools are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.
Please install the package [`gert`](https://github.com/r-lib/gert)[@R-gert] which wraps Git commands and [`repro`](https://github.com/aaronpeikert)[@R-repro] so they can act as assistance systems while you follow the tutorial:
